---
title: "AGHY Bayes model methods"
author: "Tom Miller and Marion Donald"
date: "December 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(R2jags)
library(plyr)
library(tidyverse)
library(mcmcplots)
library(gridExtra)
library(grid)
library(xlsx)
library(lme4)

#### Functions ########
Mode <- function(x) {
     ux <- unique(x)
     ux[which.max(tabulate(match(x, ux)))]
}

invlogit<-function(x){exp(x)/(1+exp(x))}
```

## Overview
These are notes and methods from Tom and Marion's "AGHY day" on December 21, 2017, with the goal of completing and finalizing parameter estimation for the AGHY vital rates, including effects of symbiosis and precipitation treatment. We recently decided, full the sake of maximum rigor and novelty, to build a population model from the vital rate data and to incorporate seed bank dynamics into the model, using Andrew Bibian's ms data. The resulting manuscript would emphasize the careful examination of context dependent symbiont effects on hosts throughout the life cycle, including rarely studied "storage" stages, and would connect the individual-level processes to population-level patterns of symbiont prevalence. We can not only use the model to explore prevalence patterns, but we can also test model predictions because we have direct observations of how prevalence changed. 

Here is the full population model that we walked through:
$$\mathbf{A}=\begin{bmatrix} 
0 & r^{-}(1-g_0^{-}) & 0 & r^{+}(1-\tau)(1-g_0^{+})\\
d^{-}g_1^{-} & r^{-}g_0^{-}+s^{-} & s^{+}g_1^{+}(1-\phi) & r^{+}(1-\tau)g_0^{+}+s^{+}(1-\rho)\\
0 & 0 & 0 & r^{+}\tau(1-g_0^{+})\\
0 & 0 & s^{+}g_1^{+}\phi & r^{+}\tau g_0^{+}+s^{+}\rho\\
\end{bmatrix}$$

The population vector is E- seed bank, E- above-ground plants, E+ seed bank, E+ above ground plants. The vital rates are: seed production per plant ($\mathit{r}$), annual survival of a seed in the seed bank ($\mathit{d}$) or an above-ground plant ($\mathit{s}$), and germination of a 0-yo seed ($\mathit{g_0}$) or a 1-yo seed ($\mathit{g_1}$). Each of these can differ between E+ and E- plants (indicated ). The symbiont loss parameters are given by greek symbols: vertical transmission ($\mathit{\tau}$), symbiont retention in surviving plants ($\mathit{\rho}$), and symbiont survival in the seed bank ($\mathit{\phi}$). We assume that horizontal transmission never occurs, and that seeds cannot remain in the seed bank for more than one year (the bank-to-bank transition is zero). We could relax either or both of these assumptions later. We know that horizontal transmission did not occur (we have numbers for this) but the longevity of the seed bank is more of a guess. 

Marion is working on estimating the above-ground parameters in a Bayesian framework using data from the Nacogdoches plots. My job is to estimate the seed bank parameters $\mathit{d}$, $\mathit{g_1}$, and $\mathit{\phi}$. Note that $\mathit{g_0}$ is a seed process (probability that a 0-yo seed recruits as a plant) but this will come from Marion's data where we can include precipitation treatment effects. We cannot include precip. effects in the other seed bank parameters because Andrew's experiment did not include this manipulation; we will therefore assume $\mathit{d}$, $\mathit{g_1}$, and $\mathit{\phi}$ do not depend on precipitation. 

### Seed bank parameter estimation
Now for the hard part: trying to re-create Andrew's analyses from 2015. Here I am going to be relying heavily on Andrew's script 'SeedBankPopModelAGHYFinal.R', which I located in the Dropbox folder 'SeedBankBayesStats (1)/UpdatedBayesStates'. This script is time-stamped 4/29/2105, probably just before his defense. It is well commented and appears to be the most mature version of the seed bank analyses, and it corresponds well to what's in his final thesis. We need to confirm with AJB that this script is the right one to use, and I will be vetting the code below to the best of my ability. Most of the R chunks below are pasted directly from this script, and I will trim these of scraps and modify as needed.

```{r data source 1, echo=T}
###############################################
##############  Data source 1 #####################
##################################################
## Reading in Bag Data
bag<-read.csv("C:\\Users\\tm9\\Dropbox\\SeedBankBayesStats (1)\\UpdatedBayesStats\\FilledBagBankDataUse.csv")
bag<-bag[c(which(bag$checked>=0)),]
bag$year<-as.factor(bag$year)

# Making 2012 Data for only plus & minus seed types
bag2012<-bag[bag$year=="2012" & bag$status!="mix",]

# Doing this for compressing data 
bag2012.AJB <-ddply(bag2012, .(spp, time, status), summarize, seeds =sum(seeds), 
		actual_seed =sum(actual_seed), gathered=sum(gathered), checked=sum(checked), 
		dead_adjusted=sum(dead_adjusted) )	
head(bag2012.AJB)
#############################################	
	
```

At this point I am not clear on what these quantities mean, ie what is 'seeds','actual_seeds','gathered', etc. I will need to ask Andrew and/or read metadata and/or figure it out from context below. I am going to further subset the 'bag' data to only include AGHY, and will apply similar subsets to the data below. I am also going to translate the ddply operations to 'modern' tidyverse.
```{r,echo=T,eval=T}
bag2012.TEX <- bag2012 %>% 
  filter(spp=="AGHY") %>%
  group_by(time,status) %>%
  summarise(mean(seeds), actual_seed =sum(actual_seed), gathered=sum(gathered), checked=sum(checked), 
		dead_adjusted=sum(dead_adjusted))
head(bag2012.TEX)
```

Moving on to next data sources.
```{r data source 2,echo=T}
###############################################
##############  Data source 2 #####################
##################################################
# Reading in seed stock data
stock<-read.csv("C:\\Users\\tm9\\Dropbox\\SeedBankBayesStats (1)\\UpdatedBayesStats\\Filledbagt0Data.csv")
stock$year<-as.factor(stock$year)

# Taking only 2012 seed data, no mix, and subset to AGHY
stock2012<-stock[stock$year=="2012" & stock$spp=="AGHY",]
```

There is a cryptic note in data source 3 about alternative data sets that do or do not include extra scores for AGHY, which apparently influence infection frequency. Need to look into this. The name of the file used in the most up to date script suggests that it includes the extra scores. WHen I tried to read in InfectSurvDataUpdated.csv I get an error, presumably this file is not in the "updated" folder. Moving forward assuming this is ok but should double check with AJB.
```{r data source 3,echo=T}
####################################################################
# Data source 3: Ep Bank Survival & Vertical Transmission Rate #####
#####################################################################
# This is the raw data for seed scoring: Infection frequencies and survival
obs.surv<-read.csv("C:\\Users\\tm9\\Dropbox\\SeedBankBayesStats (1)\\UpdatedBayesStats\\InfectSurvDataUpdatedExtraScores.csv")
#obs.surv<-read.csv("InfectSurvDataUpdated.csv", header=T, sep=",")
# The hashed out csv has data for extra enodphyte scores for AGHY(boost infection frequency for AGHY)
# Not using right now
obs.surv$year<-as.factor(obs.surv$year)

epviab.df.AJB<- ddply(obs.surv[obs.surv$endo==1 & obs.surv$traynumber !="extra",], .(species, generation), summarize, average=(sum(viability_full)/length(viability_full)), ss=length(viability_full))

emviab.df.AJB<-ddply(obs.surv[obs.surv$endo==0  & obs.surv$traynumber !="extra",], .(species, generation), summarize, average=(sum(viability_full)/length(viability_full)), ss=length(viability_full))

```
Again, I am going to convert the ddply operations to tidyverse. I am also collapsing the two df's into one. Will probably need to modify code downstream.
```{r,echo=T}
viab.df.TEX<-obs.surv%>%
  filter(species=="AGHY",endo==1,traynumber!="extra")%>%
  group_by(generation,as.factor(endo))%>%
  summarise(average=(sum(viability_full)/length(viability_full)), 
            ss=length(viability_full))

head(viab.df.TEX)
```
Next, germination data.
```{r data source 4, echo=T}
###############################################
##############  Data source 4: Germination Trials
##################################################
germ <-read.csv("C:\\Users\\tm9\\Dropbox\\SeedBankBayesStats (1)\\UpdatedBayesStats\\GermData.csv")
##TEX:not sure why these steps from AJB:
germ$count<-rep(1, nrow(germ))
germ $months<-rep(NA, nrow(germ))

# Doing this to compress data
germ.AJB<-ddply(germ, .(plottype, species, year, gen), summarize, seeded=sum(seeded), germtotal=sum(germtotal), othertotal=sum(othertotal), controltotal=sum(germ_adj_control), ss=sum(count), germ.hat=germtotal/seeded, bank.hat=(1-germ.hat))

```
And Tom's tidyverse conversion.
```{r,echo=T}
germ.TEX<-germ%>%
  filter(species=="AGHY")%>%
  group_by(plottype, year, gen)%>%
  summarize(seeded=sum(seeded), germtotal=sum(germtotal),
            othertotal=sum(othertotal),
            controltotal=sum(germ_adj_control), 
            ss=sum(count), germ.hat=germtotal/seeded,
            bank.hat=(1-germ.hat))
head(germ.TEX)
```
Last piece we will need.
```{r data source 5, echo=T}
###############################################
##############  Data source 5: Retention Data
##################################################
ret.data<-read.csv("C:\\Users\\tm9\\Dropbox\\SeedBankBayesStats (1)\\UpdatedBayesStats\\Testing.csv")
ret.data<-ret.data[ret.data$species=="AGHY",]
#read.csv("RetData.csv", sep=",", header=T)

# Checking how much data there actually is
#
ddply(ret.data[ret.data$group == "plus",], .(generation), summarize, total=length(endo), endo=sum(endo), ret.hat= endo/total)

ret.data%>%
  filter(group=="plus")%>%
  group_by(generation)%>%
  summarize(total=length(endo), 
            endo=sum(endo), 
            ret.hat= endo/total)
```

